Which fifteen-minutes neighborhoods are dead-ends? An analysis of the network attributes of fifteen-minute pedsheds

### 2024-10-18 paper/paper.Rmd
Copying the analysis to paper

@@

"Il faut oublier la traversée de Paris d'est en ouest en voiture"^[ "We must forget about crossing Paris from east to west by car"] [@alimi2020].

With this declaration during her reelection campaign in 2020, Anne Hidalgo brought international attention to the concept of walkable fifteen-minutes neighborhoods.

>>>>

```{r load-packages}
library(broom) # Convert Statistical Objects into Tidy Tibbles
library(dplyr) # A Grammar of Data Manipulation
library(evtree) # Evolutionary Learning of Globally Optimal Trees
library(ggparty) # 'ggplot' Visualizations for the 'partykit' Package
library(ggplot2) # Create Elegant Data Visualisations Using the Grammar of Graphics
library(glue) # Interpreted String Literals
library(gt) # Easily Create Presentation-Ready Display Tables
library(rpart) # Recursive Partitioning and Regression Trees
library(rpart.plot) # Plot 'rpart' Models: An Enhanced Version of 'plot.rpart'
library(patchwork) # The Composer of Plots
library(partykit)
library(tidyr) # Tidy Messy Data
library(tree) # Classification and Regression Trees
library(sf) # Simple Features for R
library(skimr) # Compact and Flexible Summaries of Data
library(SOMbrero) # SOM Bound to Realize Euclidean and Relational Outputs
library(stringr) # Simple, Consistent Wrappers for Common String Operations
library(units) # Measurement Units for R Vectors
library(vaccHamilton) # A Data Package to Estimate Accessibility of Vaccination Sites in Hamilton, ON
```

```{r load-data}
 # From package vaccHamilton
data("urban_types")

# Analysis-ready data. See notebooks in folder data-raw for details
load(file = glue::glue(here::here(),
                       "/data/amenities_walkshed_da.rda"))
load(file = glue::glue(here::here(),
                       "/data/hamilton_net.rda"))
load(file = glue::glue(here::here(),
                       "/data/urban_hammer_da.rda"))
load(file = glue::glue(here::here(),
                       "/data/walksheds_da.rda"))
load(file = glue::glue(here::here(),
                       "/data/walksheds_da_net_vars.rda"))
```

The aim is to train a classification tree to profile urban and suburban walksheds based on the attributes of the network. We need to drop 1 observation that has an NA in `transitivity`:

```{r}
d <- walksheds_da_net_vars |>
  select(-girth) |>
  drop_na() |>
  mutate(normalized_motifs_3 = motifs_3/n_edges,
         normalized_motifs_4 = motifs_4/n_edges)
```

```{r identify-network-characteristics-by-urban-type}
#seed1 <- 908544
set.seed(908544)
ev.tree <- evtree::evtree(Type ~ ., 
                data = d |> select(-c(GeoUID, 
                                      motifs_3, 
                                      motifs_4, 
                                      n_edges, 
                                      n_nodes)))

ct.tree <- partykit::cttree.treeree(Type ~ ., 
                data = d |> select(-c(GeoUID, 
                                      motifs_3, 
                                      motifs_4, 
                                      n_edges, 
                                      n_nodes)))

rpart.tree <- rpart::rpart(Type ~ ., 
                data = d |> select(-c(GeoUID, 
                                      motifs_3, 
                                      motifs_4, 
                                      n_edges, 
                                      n_nodes)))

tree.tree <- tree::tree(Type ~ ., 
                data = d |> select(-c(GeoUID, 
                                      motifs_3, 
                                      motifs_4, n_edges,
                                      n_nodes)))

```

```{r}
evtree.cm <- data.frame(obs = d$Type,
                        pred = predict(ev.tree, newdata = d)) |>
  mutate(obs = ifelse(obs == "Urban", 1, 0),
         pred = ifelse(pred == "Urban", 1, 0))

xtabs(obs ~ pred,
      data = evtree.cm)
```

```{r}
rpart.tree |>
  as.party() |>
ggparty() +
  geom_edge() +
  geom_edge_label(size = 2.5) +
  geom_node_label(aes(label = splitvar ),
                  id = "inner",
                  size = 2, nudge_y = -0.022)
```

@@

> "Il faut oublier la traversée de Paris d'est en ouest en voiture"^[ "We must forget about crossing Paris from east to west by car"]. 
> --Anne Hidalgo

With this declaration during her reelection campaign in 2020, Anne Hidalgo brought international attention to the concept of walkable fifteen-minutes neighborhoods [@alimi2020].
